Temporal graph realization from fastest paths

### 2023-12-20 SAND2024/SAND24-longversion.tex
SAND

\begin{figure}[t]
    \centering
	\begin{tikzpicture}[xscale=1.5]
		%%%S_uv
		\node[vert,label=below:$v_1$] (1) at (1,0) {};
		\node[vert,label=below:$v_2$] (2) at (3,0) {};
		\node[vert,label=below:$v_3$] (3) at (5,0) {};
		\node[vert,label=below:$v_4$] (4) at (7,0) {};
		\node[vert,label=below:$v_5$] (5) at (9,0) {};
		\draw (1) -- node[label=above:$10t+7$] {} (2) -- node[label=above:$10t+3$] {}  (3) -- node[label=above:$10t+5$] {} (4) -- node[label=above:$10t+1$] {} (5);	
	\end{tikzpicture}
	\caption{An example of a $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$, where $\Delta = 10$ and the 10-periodic labeling $\lambda: E \rightarrow \{1,2,\ldots,10\}$ is as follows: $\lambda(v_1 v_2)=7$, $\lambda(v_2 v_3)=3$, $\lambda(v_3 v_4)=5$, and $\lambda(v_4 v_5)=1$. 
    Here, the fastest temporal path from $u$ to $v$ traverses the first edge $v_1v_2$ at time $7$, second edge $v_2v_3$ a time $13$, third edge $v_3v_4$ at time $15$ and the last edge $v_4v_5$ at time $21$.
    This results in the total duration of $15$ for the fastest temporal path from $v_1$ to $v_5$.
	\label{fig:periodic-example}}
\end{figure}

@@

\problemdef{\deltaExactLong\ (\deltaExact)}
{An $n \times n$ matrix $D$, a positive integer $\Delta$.}
{Does there exist a graph $G=(V,E)$ with vertices $\{v_1,\ldots,v_{n}\}$ 
and a $\Delta$-periodic labeling $\lambda: E \rightarrow \{1,2,\ldots,\Delta\}$ such that, 
for every $i,j$, the duration of the fastest temporal path from $v_i$ to $v_j$ in the $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$ is $D_{i,j}$?}

@@

Before presenting the structure of our algorithm for \cref{thm:FPTwrtFES}, observe that, in a static graph, the number of paths between two vertices can be upper-bounded by a function $f(k)$ of the feedback edge number $k$ of the graph. 
Therefore, for any fixed pair of vertices $u$ and $v$, we can ``guess'' the edges of the fastest temporal path from $u$ to $v$. 
However, 
% due to the difficulties illustrated in~\cref{fig:ftpExample}, 
for an FPT algorithm with respect to $k$, we cannot afford to guess the edges of the fastest temporal path for each of the $O(n^2)$ pairs of vertices. 
To overcome this difficulty, our algorithm follows this high-level strategy:
\begin{itemize}
    %%%\item We perform a preprocessing step by iteratively removing all vertices of degree one.
    \item We identify a small number $f(k)$ of ``important vertices''; these consist of the sets that we call $U, U^{\ast}, Z^{\ast}$.
    \item For each pair $u,v$ of important vertices, we guess the edges of the fastest temporal path from $u$ to $v$ (and from $v$ to $u$).
    \item From these guesses we can still not deduce the edges of the fastest temporal paths between many pairs of non-important vertices. However, as we prove, it suffices to guess only a small number of specific auxiliary structures (to be defined later).
    \item From these guesses we deduce fixed relationships between the labels of most of the edges of the graph. 
    \item For all the edges, for which we do not have deduced a label yet, we introduce a \emph{variable}. Using all these variables, we build an Integer Linear Program (ILP). 
    Among the constraints in this ILP we have that, for each of the $O(n^2)$ pairs of vertices $u,v$ in the graph, the duration of one specific temporal path from $u$ to $v$ (according to our guesses) is \emph{equal} to the desired duration $D_{u,v}$, while the duration of each of the other temporal paths from $u$ to $v$ is \emph{at least} $D_{u,v}$.
    \item By making any of the above guesses, we restrict the solution space for the problem \deltaExact. This restricted solution space coincides with the set of feasible solutions to the resulting ILP. Furthermore, the set of feasible solutions for all constructed ILPs coincide with the set of all solutions to \deltaExact\ (\ie regardless of our guesses). As each ILP can be solved in FPT time with respect to $k$ by Lenstra's Theorem~\cite{Lenstra1983Integer} (the number of variables is upper bounded by a function of $k$), we obtain our FPT algorithm for \deltaExact\ with respect to $k$.
\end{itemize}

For the remainder of this section, we fix the following notation. 
Let $D$ be the input matrix of \deltaExact\ \ie
the matrix of the fastest temporal paths between all pairs of $n$ vertices, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
With $F$ we denote a minimum feedback edge set of $G$, and with $k$ the feedback edge number of $G$.
We are now ready to present our FPT algorithm. For an easier readability we split the description and analysis of the algorithm in five subsections.
We start with a preprocessing procedure for graph $G$, where we define a set of interesting vertices which then allows us to guess the desired structures.
Next we introduce some extra properties of our problem, that we then use to create ILP instances and their constraints.
At the end we present how to solve all instances and produce the desired labeling $\lambda$ of $G$, if possible.

@@

\subsubsection{Preprocessing of the input}
From the underlying graph $G$ of $D$ we first create a graph $G'$ by
iteratively removing vertices of degree one from $G$,
and denote with 
\[
Z = V(G) \setminus V(G').
\]
Then we determine a minimum feedback edge set $F$ of $G'$.
%by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
Note that $F$ is also a minimum feedback edge set of $G$.
Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, with $F$ being the corresponding feedback edge set of $G'$.
Let $V_1 \subseteq V(G')$ be the set of leaves in the spanning tree $T$, $V_2 \subseteq V(G')$ be the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subseteq V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with 
\[U = V_1 \cup V_2 \cup V_3\]
the set of vertices of interest. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie 
\[U^* = \{v \in V(G') \setminus U  :  u \in U, v \in N(u)\}.\]
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, since every neighbor of $u$ is the first vertex of a (unique) path to another vertex in $U$.
It follows that $|U^*| \in O(k^2)$.
From the construction of $Z$ (by iteratively removing vertices of degree one from $G$)
it follows that $Z$ consists of disjoint trees $T_1, T_2, \dots$.
For a tree $T_i$ we denote with $u_i$ the vertex in $G'$ that is a neighbor of a vertex in $T_i$, 
and call it a \emph{clip vertex of the tree $T_i$}.
It follows that there can be many different trees $T_i$ that are incident to the same clip vertex $u_i \in V(G')$,
but each tree $T_i$ is incident to exactly one clip vertex $u_i \in V(G')$.
Since $u_i$ is the only vertex connecting all of the trees $T_i$ incident to it,
from now on we assume that a tree $T_{u_i}$ in $Z$ is a union of trees on vertices from $V(G) \setminus V(G')$,
that are clipped at the same vertex $u_i \in V(G')$.
For each of the trees $T_{u_i}$ in $Z$, we select one vertex $r_i$, that is a neighbor of the clip vertex $u_i$,
and call it \emph{a representative vertex of the tree $T_{u_i}$}.
We now define as $Z^*$ the set of representatives $r_i$ of trees $T_i \in Z$,
where the clip vertex $v_i$ of $T_i$ is a vertex of interest, \ie
\[
Z^* = \{ r_i  :  r_i \in T_i, \text{ where $T_i \in Z$, the clip vertex $u_i$ of $T_i$ is in $U$, and $r_i u_i \in E(G)$}\}.
\]
Since there are $O(k)$ vertices of interest, we get that $|Z^*| \in O(k)$.
Finally, the set of \emph{important vertices} is defined as the set $U \cup U^{\ast} \cup Z^{\ast}$.
For an illustration see \cref{fig:labelingVertices}. Note that determining sets $U$, $U^*$, and $Z^*$ takes linear time.

@@

\subparagraph{\boldmath Considering the paths for vertices from $Z$.}
All of the above is enough to determine the labeling $\lambda$ of $G'$. Now we have to make sure that the labeling considers also the vertices in $Z$ that we initially removed from $G$.
Remember that removed vertices form disjoint trees in $G$.
Let us denote $Z$ as the set of disjoint trees, \ie $Z = T_1 \cup T_2 \cup \cdots \cup T_t$, where $T_i$ represents one of the trees.
Since there is a unique (static) path between any two vertices $z_1, z_2$ in a tree $T_i$,
it follows that there is also a unique (therefore also the fastest) temporal path between them.
Thus determining the label of an edge in $T_i$ uniquely determines the labels on all other edges of tree $T_i$.
Let us describe now how to determine the labels on edges of an arbitrary $T_i \in Z$.
Recall that for every tree $T_i$ there is a representative vertex $v_i$ of $T_i$,
and a clip vertex $u_i \in V(G')$, such that $v_i \in N_G(u_i)$.
To determine the correct label of all edges of $T_i$ we use the following property.
\begin{lemma}
Let $T_i$ be a tree in $Z$ and let $e_i = (u_i,r_i)$ be an edge in $G$, where $u_i \in V(G')$ is a clip vertex of $T_i$ and $r_i \in T_i$ is a representative of $T_i$.
Let $v \in N_{G'}(u_i)$ be the closest vertex to $r_i$, regarding the values of $D$, \ie  $D_{r_i,v} \leq D_{r_i,w}$ for all $w \in N_{G'}(u)$.
Then the path $P^*=(r_i,u_i,v)$ has to be the fastest temporal path from $r_i$ to $v$ in $G$.
\end{lemma}
\begin{proof}
Suppose that this is not true.
Then there exists a faster path $P^*_2$ from $r_i$ to $v$, that goes through the clip vertex $u_i$ of $T_i$ (as this is the only neighbor of $r_i$), through some other vertex $w \in N_{G'}(u) \setminus \{v \}$, 
and through some other path $P'$ in $G$, before it finishes in $v$, where $P'$ is at least an edge (from $w$ to $v$). 
Therefore $P^*_2= (r_i,u_i,w,P',v)$, where $d(P^*_2) \leq d(P^*)$.
Now since $D_{r_i,v} \leq D_{r_i,w}$ for all $ w\in N_{G'}(u)$ the first part of path $P^*_2$ from $r_i$ to $w$ takes at least $D_{r_i,v}$ time.
Since $v \neq w$ we need at least one more time-step (one more edge) to traverse from $w$ to reach $v$. 
So $d(P^*_2) \geq D_{r_i,v} + 1$, and so $P^*_2$ cannot be faster than $P^*$.
\end{proof}
Suppose now that we know that 
$(r_i,u_i,v)$ is the fastest temporal path from the representative $r_i$ of $T_i$ to the vertex $v$ in the neighborhood of the clip vertex $v_i$ of $T_i$.
Then
we can determine the label of edge $r_i u_i$ as 
$\lambda (r_i u_i) \equiv \lambda(u_i v) + 1 - D_{r_i,v} \pmod \Delta$.
Now, using the algorithm for trees (see \cref{thm:deltaExact-PolyTimeTrees}), we determine labels on all edges of $T_i$.
We repeat this procedure for all trees in $Z$.
What remains, is to add the equality (resp.~inequality) constraints for the fastest (resp. non-fastest) temporal paths from vertices of $Z$ to all other vertices in $G'$ and vice versa.
Note that since 
there is a unique path between vertices of tree $T_i$, and since all edges of tree $T_i$ are determined with respect to the same label,
we present the study for cases when we find fastest temporal paths from and to the 
representative vertex $r_i$ of tree $T_i$.
Each of these paths are then uniquely extended to all vertices in $T_i$.

@@

    Let $x = z_i$ be a part of a segment $S_{w,z} = (w = z_1, z_2, \dots, z_r = z)$, and denote with $z_{i-1}$ and $z_{i+1}$ the neighbouring vertices of $x$, 

@@

Let us observe some properties of the matrix $D$ from \deltaExact, when the underlying graph $G$ of $D$ is a cycle $C_n = \{v_1, v_2, \dots, v_n\}$  on $n$ vertices.
By the definition, each vertex is on distance $0$ from itself, and therefore also the duration $0$. This corresponds with all diagonal entries of $D$ being $0$.
Now, let us observe that each vertex $v_i$ has exactly two neighbours in $C_n$, namely $v_{i-1}$ and $v_{i+1}$,
therefore for all $i\in [n]$ we set $D_{i,i-1} = D_{i, i+1} = 1$, where indices are taken modulo $n$.
This results in the upper and lower diagonal of $D$ having all $1$'s, together with $D_{1,n} = D_{n,1} = 1$.
The matrix $D$ is of the following form
\begin{equation} \label{eq:D-matrixForCycles}
  D = 
  \begin{bmatrix}
    0 & 1 &  &  &  & 1\\
    1 & 0 & 1 &  &  & \\
    & 1 & 0 & 1 & & \\
    & & \ddots & \ddots & \ddots &   \\
    &   & &1 &  0 & 1\\ 
    1 &  &  &  &1 &  0 
  \end{bmatrix},
\end{equation}
where the empty entries consists of positive integers different than $1$.

@@

    Suppose for the contradiction that the statement of the claim does not hold. Then there exists a vertex $v_i \in C_n$, for which we did not calculate its travel delay.
    Let $v_j \in C_n \setminus \{v_i\}$ be an arbitrary vertex. 
    Note that the only time we cannot calculate the travel delay at vertex $v_i$, when considering vertex $v_j$, is in this case when $v_i$ is one of the maximum elements in $D_j$, \ie $v_i$ is a vertex that is on a maximum duration from $v_j$.
    This has to hold for all vertices $v_j$, therefore $v_i$ has to be on the maximum duration from every vertex $v_j \in C_n \setminus \{v_i\}$.
    It then also has to hold for vertices $v_{i-1}, v_{i+1}$, that are neighbours of $v_i$.
    We know that in this case $d^+(i-1,i) = 1, d^-(i+1,i) = 1$.
    Since $v_i$ has to be on the maximum duration from both of them 
    (\ie is one of the two maximum values $m^1_{i-1}, m^2_{i-1}$ for vertex $v_{i-1}$ and one of the two maximum values $m^1_{i+1}, m^2_{i+1}$ for vertex $v_{i+1}$),
    we know that 
    \begin{align}\label{eq:cycle-travel-delays-all}
        & d^+(i+1,i-1) < d^-(i+1,i-1) & \text{and}\\
        & d^-(i-1,i+1) < d^+(i-1,i+1).
    \end{align}
    If this would not hold, the fastest path would go through $v_i$ and we would be able to calculate the travel delay at $v_i$.

@@

\begin{comment}
We have the following two cases.
\begin{enumerate}[(i)]
    \item The clip vertex $x$ of the tree $T_i$ is not a vertex of interest.
    Let $x$ be a part of a segment $S_{w,z} = (w, z_1, z_2, \dots, z_r, z)$, and denote with $z_{i-1}$ and $z_{i+1}$ the neighbouring vertices of $x$, 
    where $z_{i-1}$ is closer to $w$ in $S_{w,z}$ and 
    $z_{i+1}$ is closer to $z$ in $S_{w,z}$.
    Similarly as above, we determine two cases.
    \begin{itemize}
    \item $S_{w,z}$ has no not yet determined edge $p$ in the middle, or $p \cap \{x \} = \emptyset$.
    Then we can determine the waiting time at vertex $x$, when traversing from $r_i$ to $z_{i-1}$ (resp.~$z_{i+1}$).
    Let $P$ be the underlying path of a fastest temporal path from $x$ to $y$ (we determined it in one of the previous steps, when we were considering vertices $x$ and $y$).
    W.l.o.g. suppose that $P$ passes through vertex $z_{i-1}$. 
    Then we set $P^-$ as $P^- = \{r_i, x\} \cup P^-$ and check the value of $d(P^-)$.
    If $d(P^-) < D_{r_i,y}$ we stop with calculations and determine that the ILP instance hos no valid solution.
    If $d(P^-) = D_{r_i,y}$ we determine that $P^-$ is the underlying path of the fastest temporal path from $r_i$ to $y$. 
    If $d(P^-) > D_{r_i,y}$ we determine that the fastest temporal path from $r_i$ to $y$ goes through the vertex $z_{i+1}$.
    We determine $P^+$ as the union of the unique path from $x$ to $z_p$, and the underlying path of the fastest temporal path $P'$ of the form $z_p \rightarrow z \leadsto y$, \ie $P^+ = (r_i, x, z_{i+1}, \dots, z_p) \cup P'$.
    We know how to determine $P'$, since
    we did it in one of the previous steps, when we were considering fastest path from $z_p$ through $z$ to $y$.
    Note, if the fastest temporal path from $z_p$ to $y$ does not pass through $z$, then we would be in previous case, where the fastest temporal path from $r_i$ to $y$ would be $P^-$.

>>>>

\begin{figure}[t]
    \centering
	\begin{tikzpicture}[xscale=1.5]
		%%%S_uv
		\node[vert,label=below:$v_1$] (1) at (1,0) {};
		\node[vert,label=below:$v_2$] (2) at (3,0) {};
		\node[vert,label=below:$v_3$] (3) at (5,0) {};
		\node[vert,label=below:$v_4$] (4) at (7,0) {};
		\node[vert,label=below:$v_5$] (5) at (9,0) {};
		\draw (1) -- node[label=above:$10t+7$] {} (2) -- node[label=above:$10t+3$] {}  (3) -- node[label=above:$10t+5$] {} (4) -- node[label=above:$10t+1$] {} (5);	
	\end{tikzpicture}
	\caption{An example of a $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$, where $\Delta = 10$ and the 10-periodic labeling $\lambda: E \rightarrow \{1,2,\ldots,10\}$ is as follows: $\lambda(v_1 v_2)=7$, $\lambda(v_2 v_3)=3$, $\lambda(v_3 v_4)=5$, and $\lambda(v_4 v_5)=1$. 
    Here, the fastest temporal path from $v_1$ to $v_2$ traverses the first edge $v_1v_2$ at time $7$, second edge $v_2v_3$ a time $13$, third edge $v_3v_4$ at time $15$ and the last edge $v_4v_5$ at time $21$.
    This results in the total duration of $21 - 7 + 1 = 15$ for the fastest temporal path from $v_1$ to $v_5$.
	\label{fig:periodic-example}}
\end{figure}

@@

\problemdef{\deltaExactLong\ (\deltaExact)}
{An integer $n \times n$ matrix $D$, a positive integer $\Delta$.}
{Does there exist a graph $G=(V,E)$ with vertices $\{v_1,\ldots,v_{n}\}$ 
and a $\Delta$-periodic labeling $\lambda: E \rightarrow \{1,2,\ldots,\Delta\}$ such that, 
for every $i,j$, the duration of the fastest temporal path from $v_i$ to $v_j$ in the $\Delta$-periodic temporal graph $(G,\lambda,\Delta)$ is $D_{i,j}$?}

@@

Before presenting the structure of our algorithm for \cref{thm:FPTwrtFES}, observe that, in a static graph, the number of paths between two vertices can be upper-bounded by a function $f(k)$ of the feedback edge number $k$ of the graph. 
Therefore, for any fixed pair of vertices $u$ and $v$, we can ``guess'' the edges of the fastest temporal path from $u$ to $v$. 
However, 
% due to the difficulties illustrated in~\cref{fig:ftpExample}, 
for an FPT algorithm with respect to $k$, we cannot afford to guess the edges of the fastest temporal path for each of the $O(n^2)$ pairs of vertices. 
To overcome this difficulty, our algorithm follows this high-level strategy:
\begin{itemize}
    %%%\item We perform a preprocessing step by iteratively removing all vertices of degree one.
    \item We identify a small number $f(k)$ of ``important vertices''; these consist of the sets that we call $U, U^{\ast}, Z^{\ast}$.
    \item For each pair $u,v$ of important vertices, we guess the edges of the fastest temporal path from $u$ to $v$ (and from $v$ to $u$).
    \item From these guesses we can still not deduce the edges of the fastest temporal paths between many pairs of non-important vertices. However, as we prove, it suffices to guess only a small number of specific auxiliary structures (to be defined later).
    \item From these guesses we deduce fixed relationships between the labels of most of the edges of the graph. 
    \item For all the edges, for which we do not have deduced a label yet, we introduce a \emph{variable}. Using all these variables, we build an Integer Linear Program (ILP). 
    Among the constraints in this ILP we have that, for each of the $O(n^2)$ pairs of vertices $u,v$ in the graph, the duration of one specific temporal path from $u$ to $v$ (according to our guesses) is \emph{equal} to the desired duration $D_{u,v}$, while the duration of each of the other temporal paths from $u$ to $v$ is \emph{at least} $D_{u,v}$.
    \item 
    %%%By making any of the above guesses, we restrict the solution space for the problem \deltaExact. This restricted solution space coincides with the set of feasible solutions to the resulting ILP. Furthermore, the set of feasible solutions for all constructed ILPs coincide with the set of all solutions to \deltaExact\ (\ie regardless of our guesses). 
    By making each of the above combinations of guesses, we essentially enumerate all possible ways that our instance of \deltaExact\ has a solution, and for each of these possible ways we create an ILP. That is, our instance of \deltaExact\ has a solution if and only if at least one of these ILPs has a feasible solution.
    As each ILP can be solved in FPT time with respect to $k$ by Lenstra's Theorem~\cite{Lenstra1983Integer} (the number of variables is upper bounded by a function of $k$), we obtain our FPT algorithm for \deltaExact\ with respect to $k$.
\end{itemize}

For the remainder of this section, we fix the following notation. 
Let $D$ be the input matrix of \deltaExact, \ie
the matrix of the fastest temporal paths between all pairs of $n$ vertices, and let $G$ be its underlying graph, on $n$ vertices and $m$ edges.
With $F$ we denote a minimum feedback edge set of $G$, and with $k$ the feedback edge number of $G$.
We are now ready to present our FPT algorithm. For an easier readability we split the description and analysis of the algorithm in five subsections.
We start with a preprocessing procedure for graph $G$, where we define a set of interesting vertices which then allows us to guess the desired structures.
Next we introduce some extra properties of our problem, that we then use to create ILP instances and their constraints.
At the end we present how to solve all instances and produce the desired labeling $\lambda$ of $G$, if possible.

@@

\subsubsection{Preprocessing of the input}
From the underlying graph $G$ of $D$ we first create a graph $G'$ by
iteratively removing vertices of degree one from $G$,
and denote with 
\[
Z = V(G) \setminus V(G').
\]
Then we determine a minimum feedback edge set $F$ of $G'$.
%by finding a spanning tree $T$ of $G'$ and set $F = E(G) \setminus E(T)$. 
Note that $F$ is also a minimum feedback edge set of $G$.
Lastly, we determine sets $U$, of \emph{vertices of interest}, and $U^*$ of the neighbors of vertices of interest, in the following way.
Let $T$ be a spanning tree of $G'$, with $F$ being the corresponding feedback edge set of $G'$.
Let $V_1 \subseteq V(G')$ be the set of leaves in the spanning tree $T$, $V_2 \subseteq V(G')$ be the set of vertices of degree two in $T$, that are incident to at least one edge in $F$, 
and let $V_3 \subseteq V(G')$ be the set of vertices of degree at least $3$ in $T$. 
Then $|V_1| + |V_2| \leq 2k$, since every leaf in $T$ and every vertex in $V_2$ is incident to at least one edge in $F$,
and $|V_3| \leq |V_1|$ by the properties of trees.
We denote with 
\[U = V_1 \cup V_2 \cup V_3\]
the set of vertices of interest. It follows that $|U| \leq 4k$.
We set $U^*$ to be the set of vertices in $V(G') \setminus U$ that are neighbors of vertices in $U$, \ie 
\[U^* = \{v \in V(G') \setminus U  :  u \in U, v \in N(u)\}.\]
Again, using the tree structure, we get that for any $u \in U$ its neighborhood is of size $|N(u)| \in O(k)$, since every neighbor of $u$ is the first vertex of a (unique) path to another vertex in $U$.
It follows that $|U^*| \in O(k^2)$.

From the construction of $Z$ (\ie by exhaustively removing vertices of degree one from~$G$), it follows that $Z$ consists of disjoint trees. 
Each of these trees has a unique neighbor $u_i$ in $G'$. Denote by $T_i$ the tree that is obtained by such a vertex $u_i$ of $G'$ and all the trees of $Z$ that are incident to $u_i$. Then  we call $u_i$ the \emph{clip vertex} of the tree $T_i$. 
For each of these trees~$T_{i}$, we select an arbitrary vertex $r_i \in Z$ that is a neighbor of the clip vertex $u_i$ in $T_i$,
and call it the \emph{representative vertex} of the tree $T_{u_i}$. 
We now define by $Z^*$ the set of all representatives $r_i$ of the trees $T_i$,
where the clip vertex $u_i$ of $T_i$ is a vertex of interest, \ie
\[
Z^* = \{ r_i  :  r_i \in Z \text{ is a neighbor of the clip vertex $u_i$ in the tree $T_i$, and $u_i \in U$}\}.
\]
Since there are $O(k)$ vertices of interest, we get that $|Z^*| \in O(k)$.
Finally, the set of \emph{important vertices} is defined as the set $U \cup U^{\ast} \cup Z^{\ast}$.
For an illustration see \cref{fig:labelingVertices}. Note that determining sets $U$, $U^*$, and $Z^*$ takes linear time.

@@

\subparagraph{\boldmath Considering the paths for vertices from $Z$.}
All of the above is enough to determine the labeling $\lambda$ of $G'$. Now we have to make sure that the labeling considers also the vertices of~$Z= V(G) \setminus V(G')$ that we initially removed from $G$.
Recall that $Z$ consists of disjoint trees, and that each of these trees has a unique neighbor $u_i$ in $G'$. 
Furthermore recall that $T_i$ denotes the tree that is obtained by such a vertex $u_i$ of $G'$ and all the trees of $Z$ that are incident to $u_i$; then $u_i$ is the \emph{clip vertex} of the tree $T_i$. 
Moreover, for each of these trees~$T_{i}$ we have selected an arbitrary vertex $r_i \in Z$ that is a neighbor of the clip vertex $u_i$ in $T_i$, 
and $r_i$ is called the \emph{representative vertex} of the tree $T_{u_i}$. 
Let us now describe now how to determine the labels on the edges of any such tree $T_i$.
To determine the correct label of all edges of $T_i$ we use the following property.
\begin{lemma}
Let $T_i$ be a tree having $u_i$ as its clip vertex, and let $e_i = (u_i,r_i)$. 
Among all neighbors of $u_i$ in $G'$, let $v \in N_{G'}(u_i)$ be the one that is closest vertex to $r_i$ (with respect to the values of $D$), \ie  $D_{r_i,v} \leq D_{r_i,w}$ for all $w \in N_{G'}(u_i)$.
Then the path $P^*=(r_i,u_i,v)$ is the fastest temporal path from $r_i$ to $v$.
\end{lemma}
\begin{proof}
Suppose that this is not true.
Then there exists a faster path $P^*_2$ from $r_i$ to $v$, that goes through the clip vertex $u_i$ of $T_i$ (as this is the only neighbor of $r_i$), through some other vertex $w \in N_{G'}(u) \setminus \{v \}$, 
and through some other path $P'$ in $G$, before it finishes in $v$, where $P'$ is at least an edge (from $w$ to $v$). 
Therefore $P^*_2= (r_i,u_i,w,P',v)$, where $d(P^*_2) \leq d(P^*)$.
Now since $D_{r_i,v} \leq D_{r_i,w}$ for all $ w\in N_{G'}(u)$ the first part of path $P^*_2$ from $r_i$ to $w$ takes at least $D_{r_i,v}$ time.
Since $v \neq w$ we need at least one more time-step (one more edge) to traverse from $w$ to reach $v$. 
So $d(P^*_2) \geq D_{r_i,v} + 1$, and so $P^*_2$ cannot be faster than $P^*$.
\end{proof}
Suppose now that we know that 
$(r_i,u_i,v)$ is the fastest temporal path from the representative $r_i$ of $T_i$ to the vertex $v$ in the neighborhood of the clip vertex $v_i$ of $T_i$.
Then
we can determine the label of edge $r_i u_i$ as 
$\lambda (r_i u_i) \equiv \lambda(u_i v) + 1 - D_{r_i,v} \pmod \Delta$.
Now, using the algorithm for trees (see \cref{thm:deltaExact-PolyTimeTrees}), we determine labels on all edges of $T_i$.
We repeat this procedure for all trees in $Z$.
What remains, is to add the equality (resp.~inequality) constraints for the fastest (resp. non-fastest) temporal paths from vertices of $Z$ to all other vertices in $G'$ and vice versa.
Note that since 
there is a unique path between vertices of tree $T_i$, and since all edges of tree $T_i$ are determined with respect to the same label,
we present the study for cases when we find fastest temporal paths from and to the 
representative vertex $r_i$ of tree $T_i$.
Each of these paths are then uniquely extended to all vertices in $T_i$.

@@

    Let $x = z_i$ be a part of a segment $S_{w,z} = (w = z_1, z_2, \dots, z_r = z)$, and denote with $z_{i-1}$ and $z_{i+1}$ the neighboring vertices of $x$, 

@@

Let us observe some properties of the matrix $D$ from \deltaExact, when the underlying graph $G$ of $D$ is a cycle $C_n = \{v_1, v_2, \dots, v_n\}$  on $n$ vertices.
By the definition, each vertex is on distance $0$ from itself, and therefore also the duration $0$. This corresponds with all diagonal entries of $D$ being $0$.
Now, let us observe that each vertex $v_i$ has exactly two neighbors in $C_n$, namely $v_{i-1}$ and $v_{i+1}$,
therefore for all $i\in [n]$ we set $D_{i,i-1} = D_{i, i+1} = 1$, where indices are taken modulo $n$.
This results in the upper and lower diagonal of $D$ having all $1$'s, together with $D_{1,n} = D_{n,1} = 1$.
The matrix $D$ is of the following form
\begin{equation} \label{eq:D-matrixForCycles}
  D = 
  \begin{bmatrix}
    0 & 1 &  &  &  & 1\\
    1 & 0 & 1 &  &  & \\
    & 1 & 0 & 1 & & \\
    & & \ddots & \ddots & \ddots &   \\
    &   & &1 &  0 & 1\\ 
    1 &  &  &  &1 &  0 
  \end{bmatrix},
\end{equation}
where the empty entries consists of positive integers different than $1$.

@@

    Suppose for the contradiction that the statement of the claim does not hold. Then there exists a vertex $v_i \in C_n$, for which we did not calculate its travel delay.
    Let $v_j \in C_n \setminus \{v_i\}$ be an arbitrary vertex. 
    Note that the only time we cannot calculate the travel delay at vertex $v_i$, when considering vertex $v_j$, is in this case when $v_i$ is one of the maximum elements in $D_j$, \ie $v_i$ is a vertex that is on a maximum duration from $v_j$.
    This has to hold for all vertices $v_j$, therefore $v_i$ has to be on the maximum duration from every vertex $v_j \in C_n \setminus \{v_i\}$.
    It then also has to hold for vertices $v_{i-1}, v_{i+1}$, that are neighbors of $v_i$.
    We know that in this case $d^+(i-1,i) = 1, d^-(i+1,i) = 1$.
    Since $v_i$ has to be on the maximum duration from both of them 
    (\ie is one of the two maximum values $m^1_{i-1}, m^2_{i-1}$ for vertex $v_{i-1}$ and one of the two maximum values $m^1_{i+1}, m^2_{i+1}$ for vertex $v_{i+1}$),
    we know that 
    \begin{align}\label{eq:cycle-travel-delays-all}
        & d^+(i+1,i-1) < d^-(i+1,i-1) & \text{and}\\
        & d^-(i-1,i+1) < d^+(i-1,i+1).
    \end{align}
    If this would not hold, the fastest path would go through $v_i$ and we would be able to calculate the travel delay at $v_i$.

@@

\begin{comment}
We have the following two cases.
\begin{enumerate}[(i)]
    \item The clip vertex $x$ of the tree $T_i$ is not a vertex of interest.
    Let $x$ be a part of a segment $S_{w,z} = (w, z_1, z_2, \dots, z_r, z)$, and denote with $z_{i-1}$ and $z_{i+1}$ the neighboring vertices of $x$, 
    where $z_{i-1}$ is closer to $w$ in $S_{w,z}$ and 
    $z_{i+1}$ is closer to $z$ in $S_{w,z}$.
    Similarly as above, we determine two cases.
    \begin{itemize}
    \item $S_{w,z}$ has no not yet determined edge $p$ in the middle, or $p \cap \{x \} = \emptyset$.
    Then we can determine the waiting time at vertex $x$, when traversing from $r_i$ to $z_{i-1}$ (resp.~$z_{i+1}$).
    Let $P$ be the underlying path of a fastest temporal path from $x$ to $y$ (we determined it in one of the previous steps, when we were considering vertices $x$ and $y$).
    W.l.o.g. suppose that $P$ passes through vertex $z_{i-1}$. 
    Then we set $P^-$ as $P^- = \{r_i, x\} \cup P^-$ and check the value of $d(P^-)$.
    If $d(P^-) < D_{r_i,y}$ we stop with calculations and determine that the ILP instance hos no valid solution.
    If $d(P^-) = D_{r_i,y}$ we determine that $P^-$ is the underlying path of the fastest temporal path from $r_i$ to $y$. 
    If $d(P^-) > D_{r_i,y}$ we determine that the fastest temporal path from $r_i$ to $y$ goes through the vertex $z_{i+1}$.
    We determine $P^+$ as the union of the unique path from $x$ to $z_p$, and the underlying path of the fastest temporal path $P'$ of the form $z_p \rightarrow z \leadsto y$, \ie $P^+ = (r_i, x, z_{i+1}, \dots, z_p) \cup P'$.
    We know how to determine $P'$, since
    we did it in one of the previous steps, when we were considering fastest path from $z_p$ through $z$ to $y$.
    Note, if the fastest temporal path from $z_p$ to $y$ does not pass through $z$, then we would be in previous case, where the fastest temporal path from $r_i$ to $y$ would be $P^-$.
